Screening Sinkhorn Algorithm for Regularized Optimal Transport

### 2019-09-25 final_NeurIPS/appendix.tex
clean comment texts ion the whole paper

% \paragraph{Bounds of}$\norm{{u}^{\textrm{sc}} - {u}^\star}_\infty $ and $\norm{{v}^{\text{sc}} - {v}^\star}_\infty $.

@@

\max\{\norm{u^\star - u^{\text{sc}}}_\infty, \norm{v^\star - v^{\text{sc}}}_\infty\} %&\leq \max\{\norm{u^\star - u^{\text{sc}}_0}_\infty +  \log(\frac{n\varepsilon^2}{c_{\mu\nu}}), \norm{v^\star - v^{\text{sc}}_0}_\infty + \log(\frac{m\varepsilon^2}{c_{\mu\nu}})\}\nonumber\\

@@

%where we used $u^{\text{sc}}_0 = \mathbf 0_n$ and $v^{\text{sc}}_0 = \mathbf 0_m.$ 

@@

\begin{align*} %\max_{i \in I_{\varepsilon,\kappa}} \mu_i %\max_{j \in J_{\varepsilon,\kappa}} \nu_j

@@

% Moreover, we have 
% \begin{align*}
% \omega_{\kappa} := (1- \kappa)\norm{{u}^{\text{sc}}}_\infty\norm{{\mu}^{\text{sc}}}_1 + (1- \kappa^{-1})\norm{{v}^{\text{sc}}}_\infty\norm{{\nu}^{\text{sc}}}_1
% \end{align*}

@@

% Setting $\omega_\kappa := (1- \kappa)\norm{\mu^{\text{sc}}}_1 + (\kappa^{-1} - 1)\norm{\nu^{\text{sc}}}_1 + \kappa^{-1} - \kappa,$ 

@@

\begin{proof}
Using inequality~\eqref{bound_on_v}, we obtain 
\begin{align*}
\norm{\mu^{\text{sc}}}_1 &= \sum_{i \in I_{\varepsilon,\kappa}} \mu^{\text{sc}}_i +  \sum_{i \in I^\complement_{\varepsilon,\kappa}}\mu^{\text{sc}}_i\\
& \overset{\eqref{i-th-marginal-mu}}{=} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + \frac \varepsilon\kappa \sum_{i \in I^\complement}\Big( \sum_{j \in J_{\varepsilon,\kappa}} K_{ij} e^{v^{\text{sc}}_j} + \varepsilon\kappa \sum_{j\in J^\complement_{\varepsilon,\kappa}}K_{ij}\Big)\\
& \overset{\eqref{bound_on_v}}{\leq} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + (n-n_b) \Big(\frac{m_b\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\kappa K_{\min}} + (m-m_b)\varepsilon^2 \Big).
% &\leq n_b\kappa + (n-n_b) \Big(\frac{m_b\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\kappa K_{\min}} + (m-m_b)\varepsilon^2 \Big)
\end{align*}
Using Remark~\ref{rem:orders_of_epsilonappa}, we get the desired closed form in~\eqref{l1-norm-mu-sc}.
Similarly, we can prove the same statement for $\norm{\nu^{\text{sc}}}_1$.
\end{proof}

@@

\LinesNotNumbered
\begin{algorithm}[tbp]
\SetNlSty{textbf}{}{.}
\DontPrintSemicolon
\caption{\textsc{Restricted \textsc{Sinkhorn}}$(K,\mu, \nu, u, v)$}
\label{restricted_sinkhorn}
\nl \textbf{set: } $\bar{f}_u = \varepsilon\kappa\, c(K_{I_{\varepsilon, \kappa}, J^\complement_{\varepsilon, \kappa}}), \bar{f}_v  = \varepsilon\kappa^{-1} \,r(K_{I^\complement_{\varepsilon, \kappa}, J_{\varepsilon, \kappa}});$\\
% \nl $\bar{f}_u = \varepsilon\kappa\, c(K_{I_{\varepsilon, \kappa}, J^\complement_{\varepsilon, \kappa}});$\\
% \nl $\bar{f}_v  = \varepsilon\kappa^{-1} \,r(K_{I^\complement_{\varepsilon, \kappa}, J_{\varepsilon, \kappa}});$\\
\nl \For{$t=1,2,3$}{
	$f^{(t)}_v \gets K^\top_{I_{\varepsilon, \kappa}, J_{\varepsilon, \kappa}} u + \bar{f}_v;$\\
	$v^{(t)} \gets \frac{\nu_{J_{\varepsilon, \kappa}}}{\kappa f^{(t)}_v} ;$\\% / (\kappa K{IJ}_v);$\\
	$f^{(t)}_u \gets K_{I_{\varepsilon, \kappa}, J_{\varepsilon, \kappa}} v + \bar{f}_u;$\\
	$u^{(t)} \gets \frac{\kappa \mu_{I_{\varepsilon, \kappa}}}{f^{(t)}_u};$\\
	$u \gets u^{(t)}$, $v \gets v^{(t)}$;
}
\nl \Return{$(u^{(t)}, v^{(t)})$}
\end{algorithm}

@@

\begin{figure*}[htbp]
	\centering
	%~\hfill~	\includegraphics[width=6.cm]{./figs/da_toy_problem.pdf}~\hfill~\\
	~\hfill~\includegraphics[width=6.cm]{./figs/wda_accur_toy.pdf}~\hfill~
	\includegraphics[width=6.cm]{./figs/wda_accur_mnist.pdf}~\hfill~\\
	\caption{Accuracy of a $1$-nearest-neighbor after WDA for the (left) toy problem and, (right) MNIST). We note a slight loss of performance for the toy problem, whereas for MNIST, all approaches yield the same performance. 
		\label{fig:wda_gain}}
\end{figure*}

\begin{figure*}[htbp]
	\centering
	%~\hfill~	\includegraphics[width=6.cm]{./figs/da_toy_problem.pdf}~\hfill~\\
	~\hfill~\includegraphics[width=6.cm]{./figs/da_toy_problem.pdf}~\hfill~
	\includegraphics[width=6.cm]{./figs/da_accur_toy_regcl100.pdf}~\hfill~\\
	~\hfill~\includegraphics[width=6.cm]{./figs/da_time_toy_regcl100.pdf}
	~\hfill~\includegraphics[width=6.cm]{./figs/da_gain_toy_regcl100.pdf}~\hfill~\\
	\caption{OT Domain Adaptation on a 3-class Gaussian toy problem. (top-left) Examples of source and target samples. (top-right) Evolution of the accuracy of a 1-nearest-neighbor classifier with respect to the number of samples. (bottom-left) Running time of the \textsc{Sinkhorn} and \textsc{Screenkhorn} for different decimation
	factors. (bottom-right). Gain in computation time. This toy problem is a problem in which classes are overlapping and distance between samples are rather limited. According to our analysis, this may be a situation in which \textsc{Screenkhorn} may
	result in smaller computational gain. We can remark that with respect to the accuracy
	\textsc{Screenkhorn} with decimation factors up to $10$ are competitive with \textsc{Sinkhorn}, although a slight loss of performance. Regarding computational time, for this example, small decimation factors does not  result in gain. However for above  $5$-factor decimation, the gain goes from  $2$ to $10$ depending on
	the number of samples. 
	\label{fig:otda:extra}}
\end{figure*}

>>>>

@@

\max\{\norm{u^\star - u^{\text{sc}}}_\infty, \norm{v^\star - v^{\text{sc}}}_\infty\} 

@@

@@

@@

@@

@@

\begin{proof}
Using inequality~\eqref{bound_on_v}, we obtain 
\begin{align*}
\norm{\mu^{\text{sc}}}_1 &= \sum_{i \in I_{\varepsilon,\kappa}} \mu^{\text{sc}}_i +  \sum_{i \in I^\complement_{\varepsilon,\kappa}}\mu^{\text{sc}}_i\\
& \overset{\eqref{i-th-marginal-mu}}{=} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + \frac \varepsilon\kappa \sum_{i \in I^\complement}\Big( \sum_{j \in J_{\varepsilon,\kappa}} K_{ij} e^{v^{\text{sc}}_j} + \varepsilon\kappa \sum_{j\in J^\complement_{\varepsilon,\kappa}}K_{ij}\Big)\\
& \overset{\eqref{bound_on_v}}{\leq} \kappa \norm{\mu_{I_{\varepsilon,\kappa}}^{\text{sc}}}_1 + (n-n_b) \Big(\frac{m_b\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\kappa K_{\min}} + (m-m_b)\varepsilon^2 \Big).
\end{align*}
Using Remark~\ref{rem:orders_of_epsilonappa}, we get the desired closed form in~\eqref{l1-norm-mu-sc}.
Similarly, we can prove the same statement for $\norm{\nu^{\text{sc}}}_1$.
\end{proof}

@@

\LinesNotNumbered
\begin{algorithm}[htbp]
\SetNlSty{textbf}{}{.}
\DontPrintSemicolon
\caption{\textsc{Restricted \textsc{Sinkhorn}}$(K,\mu, \nu, u, v)$}
\label{restricted_sinkhorn}
\nl \textbf{set: } $\bar{f}_u = \varepsilon\kappa\, c(K_{I_{\varepsilon, \kappa}, J^\complement_{\varepsilon, \kappa}}), \bar{f}_v  = \varepsilon\kappa^{-1} \,r(K_{I^\complement_{\varepsilon, \kappa}, J_{\varepsilon, \kappa}});$\\
\nl \For{$t=1,2,3$}{
	$f^{(t)}_v \gets K^\top_{I_{\varepsilon, \kappa}, J_{\varepsilon, \kappa}} u + \bar{f}_v;$\\
	$v^{(t)} \gets \frac{\nu_{J_{\varepsilon, \kappa}}}{\kappa f^{(t)}_v} ;$\\% / (\kappa K{IJ}_v);$\\
	$f^{(t)}_u \gets K_{I_{\varepsilon, \kappa}, J_{\varepsilon, \kappa}} v + \bar{f}_u;$\\
	$u^{(t)} \gets \frac{\kappa \mu_{I_{\varepsilon, \kappa}}}{f^{(t)}_u};$\\
	$u \gets u^{(t)}$, $v \gets v^{(t)}$;
}
\nl \Return{$(u^{(t)}, v^{(t)})$}
\end{algorithm}

@@

\begin{figure*}[htbp]
	\centering
	~\hfill~\includegraphics[width=6.cm]{./figs/wda_accur_toy.pdf}~\hfill~
	\includegraphics[width=6.cm]{./figs/wda_accur_mnist.pdf}~\hfill~\\
	\caption{Accuracy of a $1$-nearest-neighbor after WDA for the (left) toy problem and, (right) MNIST). We note a slight loss of performance for the toy problem, whereas for MNIST, all approaches yield the same performance. 
		\label{fig:wda_gain}}
\end{figure*}

\begin{figure*}[htbp]
	\centering
	~\hfill~\includegraphics[width=6.cm]{./figs/da_toy_problem.pdf}~\hfill~
	\includegraphics[width=6.cm]{./figs/da_accur_toy_regcl100.pdf}~\hfill~\\
	~\hfill~\includegraphics[width=6.cm]{./figs/da_time_toy_regcl100.pdf}
	~\hfill~\includegraphics[width=6.cm]{./figs/da_gain_toy_regcl100.pdf}~\hfill~\\
	\caption{OT Domain Adaptation on a 3-class Gaussian toy problem. (top-left) Examples of source and target samples. (top-right) Evolution of the accuracy of a 1-nearest-neighbor classifier with respect to the number of samples. (bottom-left) Running time of the \textsc{Sinkhorn} and \textsc{Screenkhorn} for different decimation
	factors. (bottom-right). Gain in computation time. This toy problem is a problem in which classes are overlapping and distance between samples are rather limited. According to our analysis, this may be a situation in which \textsc{Screenkhorn} may
	result in smaller computational gain. We can remark that with respect to the accuracy
	\textsc{Screenkhorn} with decimation factors up to $10$ are competitive with \textsc{Sinkhorn}, although a slight loss of performance. Regarding computational time, for this example, small decimation factors does not  result in gain. However for above  $5$-factor decimation, the gain goes from  $2$ to $10$ depending on
	the number of samples. 
	\label{fig:otda:extra}}
\end{figure*}

### 2019-09-25 final_NeurIPS/main.tex
clean comment texts ion the whole paper

\newsavebox{\tempbox}

\makeatletter
\let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}% Remove line number for one line
\makeatother

\let\oldalgorithm\algorithm
\let\oldendalgorithm\endalgorithm

\newcommand\SetAlgoFig{
  \let\algorithm\figure
  \let\endalgorithm\endfigure
}

\newcommand\SetAlgoNoFig{
  \let\algorithm\oldalgorithm
  \let\endalgorithm\oldendalgorithm
}
%%

@@

% small pmatrix
% \usepackage{amsmath}
% \newenvironment{psmallmatrix}
%   {\left(\begin{smallmatrix}}
%   {\end{smallmatrix}\right)}
%--------------------------------------------------------------------------------------------------------------------------------%

@@

\section{Conclusion}
The paper introduces a novel efficient approximation of the Sinkhorn divergence
based on a screening strategy. Screening some of the Sinkhorn dual variables
has been made possible by defining a novel constrained dual problem and by 
carefully analyzing its optimality conditions. From the latter, we derived some
sufficient conditions depending on the ground cost matrix, that some dual variables are smaller than a given threshold. Hence, we need just to solve a restricted
dual Sinkhorn problem using an off-the-shelf L-BFGS-B algorithm. We also provide
some theoretical guarantees of the quality of the approximation with respect to
the number of variables that have been screened. Numerical experiments show 
the behaviour of our \textsc{Screenkhorn} algorithm and computational time gain it can
achieve when integrated in some complex machine learning pipelines.

@@

\small
\bibliography{biblio}
\bibliographystyle{plain}
%well as with those for harvard, apalike, chicago, astron, authordate, and of course natbib.

>>>>

@@

@@

\section{Conclusion}

The paper introduces a novel efficient approximation of the Sinkhorn divergence
based on a screening strategy. Screening some of the Sinkhorn dual variables
has been made possible by defining a novel constrained dual problem and by 
carefully analyzing its optimality conditions. From the latter, we derived some
sufficient conditions depending on the ground cost matrix, that some dual variables are smaller than a given threshold. Hence, we need just to solve a restricted
dual Sinkhorn problem using an off-the-shelf L-BFGS-B algorithm. We also provide
some theoretical guarantees of the quality of the approximation with respect to
the number of variables that have been screened. Numerical experiments show 
the behaviour of our \textsc{Screenkhorn} algorithm and computational time gain it can
achieve when integrated in some complex machine learning pipelines.

@@

\small
\bibliography{biblio}
\bibliographystyle{plain}

### 2019-09-25 final_NeurIPS/section_1_introduction.tex
clean comment texts ion the whole paper

Computing optimal transport (OT) distances between pairs of probability measures or histograms, such as the earth mover's distance~\citep{werman1985,Rubner2000} and Monge-Kantorovich or Wasserstein distance~\citep{villani09optimal}, are currently generating an increasing attraction in different machine learning tasks~\citep{pmlr-v32-solomon14,kusnerb2015,pmlr-v70-arjovsky17a,ho2017}, statistics~\citep{frogner2015nips,panaretos2016,ebert2017ConstructionON,bigot2017,flamary2018WDA}, and computer vision~\citep{bonnel2011,Rubner2000,solomon2015}, among other applications~\citep{klouri17,peyre2019COTnowpublisher}.
In many of these problems, OT exploits the geometric features of the objects at hand in the underlying spaces to be leveraged in comparing probability measures.
This effectively leads to improve performance of methods that are oblivious to the geometry, for example the chi-squared distances or the Kullback-Leibler divergence.
Unfortunately, this advantage comes at the price of an enormous computational cost of solving the OT problem, that can be prohibitive in large scale applications.
For instance, the OT between two histograms with supports of equal size $n$ can be formulated as a linear programming problem that requires generally super $\bigO(n^3)$~\citep{pele2009} arithmetic operations, which is problematic when $n$ is larger than $10^3.$

A remedy to the heavy computation burden of OT lies in a prevalent approach referred as regularized OT~\citep{cuturinips13} and operates by adding an entropic regularization penalty to the original problem.  
Such a regularization guarantees a unique solution, since the objective function is strongly convex, and a greater computational stability.
%Furthermore, the dual of the entropic problem, referred to as {dual of Sinkhorn divergence}, was studied in~\cite{cuturinips13} and noticed that finding the dual solution was equivalent to finding two diagonal matrices that made a full matrix bistochastic.
More importantly, this regularized OT can be solved efficiently with celebrated matrix scaling algorithms, such as Sinkhorn's fixed point iteration method~\citep{sinkhorn1967,knight2008,kalantari2008}.

%Sinkhorn scaling for computing OT distances is a well studied problem in many recent works. 
%The main idea is to improve the matrix-vector operations that are the computational bottleneck of Sinkhorn’s algorithm. 
Several works have considered further improvements in the resolution of this regularized OT problem.
A greedy version of Sinkhorn, called Greenkhorn algorithm and proposed in~\cite{altschulernips17}, allows to select columns and rows to be updated that most violate the constraints.                   
Another approach based on low-rank approximation of the cost matrix using Nystrom method induces the Nys-Sink algorithm~\citep{altschuler2018Nystrom}. 
Other classical optimization algorithms have been considered to approximate the OT, for instance accelerated gradient descent~\citep{xie2018proxpointOT,dvurechensky18aICML,lin2019}, quasi Newton methods~\citep{blondel2018ICML,cuturi2016SIAM} and stochastic gradient descent~\citep{genevay2016stochOT,khalilabid2018}.

@@

	\item From a methodological point of view, we propose a reformulation of the dual of the Sinkhorn divergence problem by imposing variables to be larger than a threshold.
	This formulation allows us to introduce sufficient conditions, computable beforehand, for a variable to  strictly satisfy its constraint, leading then to
	a ``screened'' version of the dual of Sinkhorn divergence. 
	\item We provide some theoretical analysis of the solution of the ``screened''  Sinkhorn divergence, showing that its objective value and the marginal constraint satisfaction are properly controlled 	as the number of screened variables decreases.
	\item From an algorithmic standpoint, we use a constrained L-BFGS-B algorithm \citep{nocedal1980,byrd1995L-BFGS-B} but provide a careful analysis of the lower bound and upper bound of the dual	variables, resulting in a well-posed and efficient algorithm denoted as \emph{Screenkhorn}.
	\item Our empirical analysis depicts how the approach behaves in a simple Sinkhorn divergence computation context. When considered in  complex machine learning
	pipelines, we show that \emph{Screenkhorn} can lead to strong gain in efficiency
	while not compromising on accuracy.
\end{itemize}}
%	
%
%These constraints are defined through a convex set which depends on two parameters acting like threshold and scaling factor.
%We prove that dual solution of this reformulation guarantees the existence of two active indices sets for the potential variables.
%Furthermore, the active sets are both directly linked to a priori fixed number budget of points from the supports of the given discrete measures.
%We then restrict the constraints feasibility by taking into account the properties providing by the active sets to get a ``screened'' version of the dual of Sinkhorn divergence. 
%Screenkhorn algorithm developed in this paper consists of two steps; the first one is an initialization step devoted to determine the active sets while the second is a constrained box-constrained L-BFGS-B solver, a limited-memory quasi-Newton algorithm~\cite{nocedal1980,byrd1995L-BFGS-B} that relies on an estimation of the inverse of the Hessian based on gradients differences. 
%L-BFGS-B algorithm was adapted to the OT setting via the dual of Sinkhorn divergence in ~\cite{cuturi2016SIAM,blondel2018ICML}.}

The remainder of the paper is organized as follow. In Section~\ref{sec:regularized_discrete_ot} we briefly review the basic setup of regularized discrete OT. 
Section~\ref{sec:screened_dual_of_sinkhorn_divergence} contains our main contribution, that is, the \emph{Screenkhorn} algorithm. 
Section~\ref{sec:analysis_of_marginal_violations} devotes to theoretical guarantees for marginal violations of \emph{Screenkhorn}. 
In Section~\ref{sec:numerical_experiments} we present numerical results for the proposed algorithm, compared with the state-of-art Sinkhorn algorithm as implemented in~\cite{flamary2017pot}. 
The proofs of theoretical results are postponed to the supplementary material.

\emph{Notation.} For any positive matrix $T \in \R^{n\times m}$, we define its negative entropy as $H(T) = -\sum_{i,j} T_{ij} \log(T_{ij}).$
Let $r(T) = T\mathbf 1_m \in \R^n$ and $c(T) = T^\top\mathbf 1_n \in \R^m$ denote the rows and columns sums of $T$ respectively. The coordinates $r_i(T)$ and $c_j(T)$ denote the $i$-th row sum and the $j$-th column sum of $T$, respectively.
The scalar product between two matrices denotes the usual inner product, that is $\inr{T, W} = \text{tr}(T^\top W) = \sum_{i,j}T_{ij}W_{ij},$ where $T^\top$ is the transpose of $T$. 
We write $\mathbf{1}$ (resp. $\mathbf{0}$) the vector having all coordinates equal to one (resp. zero).
$\Delta(w)$ denotes the diag operator, such that if $w \in \R^n$, then $\Delta(w) = \text{diag}(w_1, \ldots, w_n)\in \R^{n\times n}$.
For a set of indices $L=\{i_1, \ldots, i_k\} \subseteq \{1, \ldots, n\}$ satisfying $i_1 < \cdots <i_k,$ we denote the complementary set of $L$ by $L^\complement = \{1, \ldots, n\} \backslash L$. We also denote $|L|$ the cardinality of $L$.
Given a vector $w \in \R^n$, we denote $w_L= (w_{i_1}, \ldots, w_{i_k})^\top \in \R^k$ and its complementary $w_{L^\complement} \in \R^{n- k}$.  The notation is similar for matrices; given another subset of indices $S = \{j_1, \ldots, j_l\} \subseteq \{1, \ldots, m\}$ with $j_1 < \cdots <j_l,$ and a matrix $T\in \R^{n\times m}$, we use $T_{(L,S)}$, to denote the submatrix of $T$, namely the rows and columns of $T_{(L,S)}$ are indexed by $L$ and $S$ respectively.
When applied to matrices and vectors,  $\odot$ and $\oslash$ (Hadamard product and division) and exponential notations refer to elementwise operators.
Given two real numbers $a$ and $b$, we write $a\vee b = \max(a,b)$ and $a\wedge b = \min(a,b).$

>>>>

Computing optimal transport (OT) distances between pairs of probability measures or histograms, such as the earth mover's distance~\citep{werman1985,Rubner2000} and Monge-Kantorovich or Wasserstein distance~\citep{villani09optimal}, are currently generating an increasing attraction in different machine learning tasks~\citep{pmlr-v32-solomon14,kusnerb2015,pmlr-v70-arjovsky17a,ho2017}, statistics~\citep{frogner2015nips,panaretos2016,ebert2017ConstructionON,bigot2017,flamary2018WDA}, and computer vision~\citep{bonnel2011,Rubner2000,solomon2015}, among other applications~\citep{klouri17,peyre2019COTnowpublisher}.
In many of these problems, OT exploits the geometric features of the objects at hand in the underlying spaces to be leveraged in comparing probability measures.
This effectively leads to improved performance of methods that are oblivious to the geometry, for example the chi-squared distances or the Kullback-Leibler divergence.
Unfortunately, this advantage comes at the price of an enormous computational cost of solving the OT problem, that can be prohibitive in large scale applications.
For instance, the OT between two histograms with supports of equal size $n$ can be formulated as a linear programming problem that requires generally super $\bigO(n^{2.5})$~\citep{leeSidford2013PathFI} arithmetic operations, which is problematic when $n$ becomes larger.

A remedy to the heavy computation burden of OT lies in a prevalent approach referred to as regularized OT~\citep{cuturinips13} and operates by adding an entropic regularization penalty to the original problem.  
Such a regularization guarantees a unique solution, since the objective function is strongly convex, and a greater computational stability.
More importantly, this regularized OT can be solved efficiently with celebrated matrix scaling algorithms, such as Sinkhorn's fixed point iteration method~\citep{sinkhorn1967,knight2008,kalantari2008}.

Several works have considered further improvements in the resolution of this regularized OT problem.
A greedy version of Sinkhorn algorithm, called Greenkhorn~\cite{altschulernips17}, allows to select and update columns and rows that most violate the polytope constraints.                   
Another approach based on low-rank approximation of the cost matrix using the Nystr\"om method induces the Nys-Sink algorithm~\citep{altschuler2018Nystrom}. 
Other classical optimization algorithms have been considered for approximating the OT, for instance accelerated gradient descent~\citep{xie2018proxpointOT,dvurechensky18aICML,lin2019}, quasi-Newton methods~\citep{blondel2018ICML,cuturi2016SIAM} and stochastic gradient descent~\citep{genevay2016stochOT,khalilabid2018}.

@@

	\item From a methodological point of view, we propose a new formulation of the dual of the Sinkhorn divergence problem by imposing variables to be larger than a threshold.
	This formulation allows us to introduce sufficient conditions, computable beforehand, for a variable to  strictly satisfy its constraint, leading then to
	a ``screened'' version of the dual of Sinkhorn divergence. 
	\item We provide some theoretical analysis of the solution of the ``screened''  Sinkhorn divergence, showing that its objective value and the marginal constraint satisfaction are properly controlled 	as the number of screened variables decreases.
	\item From an algorithmic standpoint, we use a constrained L-BFGS-B algorithm \citep{nocedal1980,byrd1995L-BFGS-B} but provide a careful analysis of the lower and upper bounds of the dual	variables, resulting in a well-posed and efficient algorithm denoted as \textsc{Screenkhorn}.
	\item Our empirical analysis depicts how the approach behaves in a simple Sinkhorn divergence computation context. When considered in  complex machine learning
	pipelines, we show that \textsc{Screenkhorn} can lead to strong gain in efficiency
	while not compromising on accuracy.
\end{itemize}}

The remainder of the paper is organized as follow. In Section~\ref{sec:regularized_discrete_ot} we briefly review the basic setup of regularized discrete OT. 
Section~\ref{sec:screened_dual_of_sinkhorn_divergence} contains our main contribution, that is, the \textsc{Screenkhorn} algorithm. 
Section~\ref{sec:analysis_of_marginal_violations} is devoted to theoretical guarantees for marginal violations of \textsc{Screenkhorn}. 
In Section~\ref{sec:numerical_experiments} we present numerical results for the proposed algorithm, compared with the state-of-art Sinkhorn algorithm as implemented in~\cite{flamary2017pot}. 
The proofs of theoretical results are postponed to the supplementary material.

\emph{Notation.} For any positive matrix $T \in \R^{n\times m}$, we define its entropy as $H(T) = -\sum_{i,j} T_{ij} \log(T_{ij}).$
Let $r(T) = T\mathbf 1_m \in \R^n$ and $c(T) = T^\top\mathbf 1_n \in \R^m$ denote the rows and columns sums of $T$ respectively. The coordinates $r_i(T)$ and $c_j(T)$ denote the $i$-th row sum and the $j$-th column sum of $T$, respectively.
The scalar product between two matrices denotes the usual inner product, that is $\inr{T, W} = \text{tr}(T^\top W) = \sum_{i,j}T_{ij}W_{ij},$ where $T^\top$ is the transpose of $T$. 
We write $\mathbf{1}$ (resp. $\mathbf{0}$) the vector having all coordinates equal to one (resp. zero).
$\Delta(w)$ denotes the diag operator, such that if $w \in \R^n$, then $\Delta(w) = \text{diag}(w_1, \ldots, w_n)\in \R^{n\times n}$.
For a set of indices $L=\{i_1, \ldots, i_k\} \subseteq \{1, \ldots, n\}$ satisfying $i_1 < \cdots <i_k,$ we denote the complementary set of $L$ by $L^\complement = \{1, \ldots, n\} \backslash L$. We also denote $|L|$ the cardinality of $L$.
Given a vector $w \in \R^n$, we denote $w_L= (w_{i_1}, \ldots, w_{i_k})^\top \in \R^k$ and its complementary $w_{L^\complement} \in \R^{n- k}$.  The notation is similar for matrices; given another subset of indices $S = \{j_1, \ldots, j_l\} \subseteq \{1, \ldots, m\}$ with $j_1 < \cdots <j_l,$ and a matrix $T\in \R^{n\times m}$, we use $T_{(L,S)}$, to denote the submatrix of $T$, namely the rows and columns of $T_{(L,S)}$ are indexed by $L$ and $S$ respectively.
When applied to matrices and vectors,  $\odot$ and $\oslash$ (Hadamard product and division) and exponential notations refer to elementwise operators.
Given two real numbers $a$ and $b$, we write $a\vee b = \max(a,b)$ and $a\wedge b = \min(a,b).$
